An Algorithm to Prune Integer Overflow Tests

### 2012-08-30 doc/Manuscript/CGO/paper.tex
Little corrections

%SUB instrumentation
The SUB instruction performs the subtraction of two operands $op_1$ and $op_2$ and
stores the result in a variable $x$.
The semantics of an SUB instruction is $x = op_1 - op_2$. 
Note that the SUB semantics is close to the ADD semantics. For this reason, the
instrumentation of the SUB instructions looks like the ADD instrumentation.
In this case, we insert different checks for signed and unsigned instructions, too.
If we are dealing with an unsigned SUB instruction $x$, we insert the following check:
\code{if $op_1 < op_2$ then} handle overflow \code{else} continue to 
$x$'s next instruction. Note that the assertion makes a single comparison. This is sound
because if $op_2$ is greater than $op_1$, the result would be negative. However, unsigned
integers can't represent negative numbers. In this case, we have a overflow.
The instrumentation of a signed SUB instruction $x$ will have the following assertion:
\code{if ($op_1 < 0$ and $op_2 > 0$ and $x > 0$) or ($op_1 > 0$ and $op_2 < 0$ and $x < 0$) then} 
handle overflow \code{else} continue to $x$'s next instruction.

%MUL instrumentation
The MUL instruction performs the multiplication of two operands $op_1$ and $op_2$ and
stores the result in a variable $x$.
The semantics of an SUB instruction is $x = op_1 \times{} op_2$. 
The dynamic check of the MUL instruction is the same for signed or unsigned
instructions, because it relies on DIV, the inverse operation. 
To verify if the overflow has occurred, we divide $x$ by $op_1$. If the result
is different of $op_2$, then we have an overflow to handle. 
In this case, we will have the following assertion: 
\code{if $op_1 \neq{} 0$ and $(x \div{} op_1) \neq{} op_2$ then} 
handle overflow \code{else} continue to $x$'s next instruction.
Note that if $op_1 = 0$, the result of the multiplication will never overflow,
because the result of the MUL will be $0$, too. In this case, if $op_1 \eq{} 0$,
we don't perform the DIV, to prevent a division by zero exception.
This technique of using the inverse operation to check for overflows doesn't work
for ADD and SUB instructions because the $n$-bits two's complement wrapping-arithmetics 
would give us false negatives in the assertions.

%SHL instrumentation
The SHL instruction performs the shift left operation. It receives a value $op_1$
and the number of bits $n$ the SHL will shift $op_1$ by; the operation's result
is stored in a variable $x$.
The semantics os a SHL instruction is $x = op_1 \times{} 2^n$. 
The SHL instructions result in overflows when the $op_1$ is positive and 
$x$ is negative or when $op_1$ is negative and $n \neq{} 0$, because 
the SHL in negative numbers would produce undefined results that are
as harmful as integer overflows. This assertion has the same semantics for
signed and unsigned instructions.
In this case, we will have the following check: 
\code{if ($op_1 > 0$ and $x < op_1$) or ($op_1 < 0$ and $n \neq{} 0$) then} 
handle overflow \code{else} continue to $x$'s next instruction.

@@

%Overview about the instrumentation technique
Our instrumentation library works by adding new instructions to
the target programs. The instrumentation of all the instructions works
in the same way: first, the pass inserts instructions to gather the
information needed to do the assertion.
Then, the pass inserts compare instructions to generate a boolean value
that indicates if there is an overflow or not. 
After this, the pass inserts a conditional branch: 
the false clause of the branch, where the control goes when there is no 
overflow and the program, leads to the first instruction after the instruction
which has received the instrumentation.
the true clause of the branch leads to a block where we do the
overflow handling.

>>>>

%SUB instrumentation
The SUB instruction performs the subtraction of two operands $op_1$ and $op_2$ and
stores the result in a variable $x$.
The semantics of a SUB instruction is $x = op_1 - op_2$. 
Note that the SUB semantics is close to the ADD semantics. For this reason, the
instrumentation of the SUB instructions looks like the ADD instrumentation.
In this case, we insert different checks for signed and unsigned instructions, too.
If we are dealing with an unsigned SUB instruction $x$, we insert the following check:
\code{if $op_1 < op_2$ then} handle overflow \code{else} continue to 
$x$'s next instruction. Note that the assertion makes a single comparison. This is sound
because if $op_2$ is greater than $op_1$, the result would be negative. However, unsigned
integers can't represent negative numbers. In this case, we have a overflow.
The instrumentation of a signed SUB instruction $x$ will have the following assertion:
\code{if ($op_1 < 0$ and $op_2 > 0$ and $x > 0$) or ($op_1 > 0$ and $op_2 < 0$ and $x < 0$) then} 
handle overflow \code{else} continue to $x$'s next instruction.

%MUL instrumentation
The MUL instruction performs the multiplication of two operands $op_1$ and $op_2$ and
stores the result in a variable $x$.
The semantics of a MUL instruction is $x = op_1 \times{} op_2$. 
The dynamic check of the MUL instruction is the same for signed or unsigned
instructions, because it relies on DIV, the inverse operation. 
To verify if the overflow has occurred, we divide $x$ by $op_1$. If the result
is different of $op_2$, then we have an overflow to handle. 
In this case, we will have the following assertion: 
\code{if $op_1 \neq{} 0$ and $(x \div{} op_1) \neq{} op_2$ then} 
handle overflow \code{else} continue to $x$'s next instruction.
Note that if $op_1 = 0$, the result of the multiplication will never overflow,
because the result of the MUL will be $0$, too. In this case, if $op_1 = 0$,
we don't perform the DIV, to prevent a division by zero exception.
This technique of using the inverse operation to check for overflows doesn't work
for ADD and SUB instructions because the $n$-bits two's complement wrapping-arithmetics 
would give us false negatives in the assertions.

%SHL instrumentation
The SHL instruction performs the shift left operation. It receives a value $op_1$
and the number of bits $n$ the SHL will shift $op_1$ by; the operation's result
is stored in a variable $x$.
The semantics of a SHL instruction is $x = op_1 \times{} 2^n$. 
The SHL instructions result in overflows when the $op_1$ is positive and 
$x$ is negative or when $op_1$ is negative and $n \neq{} 0$, because 
the SHL in negative numbers would produce undefined results that are
as harmful as integer overflows. This assertion has the same semantics for
signed and unsigned instructions.
In this case, we will have the following check: 
\code{if ($op_1 > 0$ and $x < op_1$) or ($op_1 < 0$ and $n \neq{} 0$) then} 
handle overflow \code{else} continue to $x$'s next instruction.

@@

%Overview about the instrumentation technique
Our instrumentation library works by adding new instructions to
the target programs. The instrumentation of all the instructions works
in the same way: first, the pass inserts instructions to gather the
information needed to do the assertion.
Then, the pass inserts compare instructions to generate a boolean value
that indicates if there is an overflow or not. 
After this, the pass inserts a conditional branch: 
the false clause of the branch, where the control goes when there is no 
overflow, leads to the first instruction after the instruction
which has received the instrumentation.
the true clause of the branch leads to a block where we do the
overflow handling.
